### 2015-09-15 Manuscript/SW_early_life_telomeres_SOM_v9.Rmd
More manuscript updates - mainly results

**Table S1** Model selection table for analysis of factors affecting early life telomere length in juvenile Seychelles warblers. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. i = intercept.

@@

#Run the model
lm1 <- lmer(TLKB ~ Agemonths + TQspace + Insect + GroupSize + Sex + Condition + Helper + (1|LayYear),
            data = FlSAall,REML=FALSE)

lm1 <- standardize(lm1)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))
colnames(t1)[2:7] <- c('Sex', 'Body condition','Group size','Helper','Food av.','TQ')
getterms <- function(x)
  {
  temp <- c('i',names(which(!(is.na(x[2:7])))))
  return(paste(temp,collapse = " + "))
  }

tableS1 <- data.frame(Model = apply(t1,1,getterms),t1[,c('df','AICc','delta','weight')])
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
kable(tableS1,digits=3)

@@

```{r Figure S1,fig.width=4,fig.height=4}

temp <- dd

ddFigS1 <- ddply(dd,
                 .(Age),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

ddFigS1 <- ddFigS1[complete.cases(ddFigS1),]
ddFigS1 <- subset(ddFigS1,Age<14)

FigS1 <- ggplot(ddFigS1,
       aes(x = Age,
           y = TLKBmean)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  ylab('Telomere length (kb)') + 
  xlab('Age') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

FigS1

@@

**Figure S1** Cross-sectional telomere length (mean $\pm$ s.e.) in relation to age in Seychelles warblers.

```{r Figure S2,fig.width=4,fig.height=4}

temp <- FlSA

ddFigS2 <- ddply(temp,
                 .(LayYear),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

ddFigS2 <- ddFigS2[complete.cases(ddFigS2),]
ddFigS2$LayYear <- as.numeric(paste(ddFigS2$LayYear))

FigS2 <- ggplot(ddFigS2,
       aes(x = LayYear,
           y = TLKBmean)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  ylab('Telomere length (kb)') + 
  xlab('Lay Year') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

FigS2

@@

**Figure S2** Cross-sectional juvenile telomere length (mean $\pm$ s.e.) across birth years in Seychelles warblers

>>>>

**Table S1** Model selection table for analysis of factors affecting early life telomere length in Seychelles warbler chicks. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here.

@@

#Run the model
lm1 <- lm(cenTL ~ cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ,
            data = chickall)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))

#Sort out row/col names
nams <- colnames(t1)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'

colnames(t1) <- nams

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:6)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

@@

**Table S2** Model selection table for analysis of factors affecting early life telomere length in Seychelles warbler fledglings. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here.

```{r Table S2}

temp <- subset(FlSAall,Ageclass == 'FL')

#Run the model
lm1 <- lm(cenTL ~ cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ,
            data = temp)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))

#Sort out row/col names
nams <- colnames(t1)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'

colnames(t1) <- nams

tableS2 <- data.frame(Model = apply(t1,1,getterms,c(2:6)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS2) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS2) <- NULL
tableS2$Model[tableS2$Model == ''] <- 'Null model'
tableS2[,3:5] <- lapply(tableS2[,3:5],as.numeric)
kable(tableS2,digits=2)

@@

**Table S3** Model selection table for analysis of factors affecting early life telomere length in Seychelles warbler sub-adults Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here.

```{r Table S3}

temp <- subset(FlSAall,Ageclass == 'SA')

#Run the model
lm1 <- lm(cenTL ~ cenHelper + cenNonHelper + Sex + cenTarsus  + cenTQ,
            data = temp)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))

#Sort out row/col names
nams <- colnames(t1)
nams[nams == 'cenTarsus'] <- 'Tarsus'
nams[nams == 'cenNonHelper'] <- 'Non helpers'
nams[nams == 'SexMales'] <- 'Sex'
nams[nams == 'cenTQ'] <- 'TQ'
nams[nams == 'cenHelper'] <- 'Helpers'

colnames(t1) <- nams

tableS3 <- data.frame(Model = apply(t1,1,getterms,c(2:6)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS3) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS3) <- NULL
tableS3$Model[tableS3$Model == ''] <- 'Null model'
tableS3[,3:5] <- lapply(tableS3[,3:5],as.numeric)
kable(tableS3,digits=2)

@@

### 2015-09-15 Manuscript/SW_early_life_telomeres_tables_and_figures v9.Rmd
More manuscript updates - mainly results

**Figure 1** Telomere length and age Seychelles warblers, based on cross-sectional **(A)** and longitudinal **(B)** data. Points and error bars represent mean and 95% confidence intervals, respectively. In **B**, telomere loss is a standardised and corrected rate (see methods for details) based on differences in telomere length between the sampling event of birds as a juvnile, and the first available sampling as an adult. Values above zero indicate telomere shortening in early life.

**Figure 2** Temporal variation in early-life telomere length in the Seychelles warbler based on cross-sectional **(A,B)** and longitudinal **(C,D)** data (see Figure 1 and main text for details). Points and error bars represent mean and standard error telomere length for each summer (blue) or winter (red) field season. In **A** and **C** points are ordered temporally, while in **B** and **D** telomere length/loss for each season is plotted against island-wide food availability. Lines and shaded areas represent fitted values and 95% confidence limits from a linear regression.

**Figure 3** Factors affecting telomere length in Seychelles warbler chicks. **A** Model averaged estimates and 95% confidence intervals for all explanatory terms used in a linear model with telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; Food av. = annual food availability, TQ = territory quality). **B** Presence/absence of helpers, **C** sex and **D** tarsus length in relation to raw telomere length. Points and error bars in **B and C** are mean $\pm$ standard error telomere lengths for each group, and the line and shaded areas from **D** represent fitted values and 95% confidence limits from a linear regression.

**Figure 4** Factors affecting telomere length in Seychelles warblers fledglings **(A)** and subadults **(B)**. Model averaged estimates and 95% confidence intervals for all explanatory terms used in a linear model with  telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; Food av. = annual food availability, TQ = territory quality).

**Figure 5** Kaplan-Meier curves showing the relationship between telomere length and survival in chick (**A**) and juvenile (**B**) Seychelles warblers. Telomere length is binned into groups here for visualisation purposes only (long and short = greater than or less than median telomere length, repectively).

@@

```{r Figure 1,fig.width=8,fig.height=9}
######################
#1A
######################

@@

Fig1A <- ggplot(ddFig1A,
       aes(x = Ageclass,
           y = TLKBmean)) +
  ylab('Corrected telomere loss') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'A',x=0.5,y=6.2)

@@

Fig1B <- ggplot(ddFig1B,
       aes(x = Ageclass,
           y = TROCmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  annotate('text',label = 'B',x=0.5,y=0.0018)

@@

ddFig2A <- ddply(juv,
                 .(CatchYear,Season),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))
ddFig2A <- ddFig2A[complete.cases(ddFig2A),]
ddFig2A <- subset(ddFig2A,TLKBse < 4)

Fig2A <- ggplot(ddFig2A,
       aes(x = CatchYear,
           y = TLKBmean,
           col = Season)) +
  ylab('Telomere length') + 
  xlab('Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse),
                  position = position_dodge(width = 1)) +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'A',x=1995,y=10)

@@

temp <- juv
temp$Insect <- factor(temp$Insect)

ddFig2B <- ddply(temp,
                 .(Insect,Season),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))
ddFig2B$Insect <- as.numeric(paste(ddFig2B$Insect))
ddFig2B <- ddFig2B[complete.cases(ddFig2B),]

Fig2B <- ggplot(ddFig2B,
       aes(x = Insect,
           y = TLKBmean,
           col = Season)) +
  ylab('Telomere length') + 
  xlab('Food availability') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  stat_smooth(method = 'lm',se=T,col = 'darkgrey') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'B',x=-1.5,y=8)

@@

ddFig2C <- ddply(Loss,
                 .(CatchYear,Season),
                 summarize,
                 TROCmean = mean(TROC),
                 TROCse = se(TROC))
ddFig2C <- ddFig2C[complete.cases(ddFig2C),]
ddFig2C <- subset(ddFig2C,TROCse < 4)

Fig2C <- ggplot(ddFig2C,
       aes(x = CatchYear,
           y = TROCmean,
           col = Season)) +
  ylab('Corrected telomere loss') + 
  xlab('Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TROCmean-TROCse,ymax=TROCmean+TROCse),
                  position = position_dodge(width = 1)) +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'C',x=1995,y=0.009)

####################################
temp <- Loss
temp$Insect <- factor(temp$Insect)

ddFig2D <- ddply(temp,
                 .(Insect,Season),
                 summarise,
                 TROCmean = mean(TROC),
                 TROCse = se(TROC))
ddFig2D$Insect <- as.numeric(paste(ddFig2D$Insect))
ddFig2D <- ddFig2D[complete.cases(ddFig2D),]

Fig2D <- ggplot(ddFig2D,
       aes(x = Insect,
           y = TROCmean,
           col = Season)) +
  geom_pointrange(aes(min=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  ylab('Corrected telomere loss') + 
  xlab('Food availability') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  stat_smooth(method = 'lm', se=T, col = 'darkgrey') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'D',x=-1.5,y=0.0075)

@@

```{r Figure 3,fig.width=8,fig.height=9}

@@

Fig3A <- forestplot(table2) +
  annotate('text',label = 'A',x=0.6,y=1.2)

@@

temp <- chicks

ddFig3B <- ddply(temp,
                 .(Sex),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

Fig3B <- ggplot(ddFig3B,
       aes(x = Sex,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'B',x=0.5,y=7)

@@

ddFig3C <- ddply(temp,
                 .(HelperF),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

Fig3C <- ggplot(ddFig3C,
       aes(x = HelperF,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'C',x=0.5,y=8)

######################
#3D
######################

Fig3D <- ggplot(chicks,
                aes(x = Tarsus,
                    y = TLKB)) +
  ylab('Telomere length (kb)') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  geom_point(size=3) +
  stat_smooth(method = 'lm',col = 'black') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'D',x=19,y=15)

multiplot(Fig3A,Fig3B,Fig3C,Fig3D,layout=matrix(c(1:4), nrow=2, byrow=FALSE))

@@

```{r Figure 5,fig.height=8,fig.width=5}

s1 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = chicks)
Fig5A <- ggsurv(s1,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='A')

s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = FlSAall)
Fig5B <- ggsurv(s2,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='B')

@@

>>>>

**Figure 1** Telomere length (mean $\pm$ s.e.) and age Seychelles warblers, based on cross-sectional **(A)** and longitudinal **(B)** data. In **B**, telomere loss is a standardised and corrected rate (see methods for details) based on the difference in telomere length from the sample taken from birds in their first year of life, and the first available sampling as an adult. Values above zero indicate telomere shortening in early life.

**Figure 2** Temporal variation in early-life telomere dynamics in the Seychelles warbler based on cross-sectional **(A,B)** and longitudinal **(C,D)** data (see Fig. 1 and main text for details). Points and error bars represent mean and standard error telomere length for all birds born in each summer (blue) or winter (red) breeding season. In **A** and **C** points are ordered temporally, while in **B** and **D** telomere length/loss for each season is plotted against island-wide food availability. Lines and shaded areas represent fitted values and 95% confidence limits from a linear regression.

**Figure 3** Factors affecting telomere length in Seychelles warbler chicks. **A** Model averaged estimates and 95% confidence intervals for all explanatory terms used in a linear model with telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; TQ = territory quality). **B** Presence/absence of helpers and **c** tarsus length in relation to raw telomere length. Points and error bars in **B** are mean $\pm$ s.e., and the line and shaded areas from **C** represent fitted values and 95% confidence limits from a linear regression.

**Figure 4** Factors affecting telomere length in Seychelles warblers fledglings **(A)** and subadults **(B)**. Points and error bars represent model averaged estimates and 95% confidence intervals for all explanatory terms used in a linear model with  telomere length as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; TQ = territory quality).

**Figure 5** Temporal variation in early-life telomere dynamics and lifespan in the Seychelles warbler, based on cross-sectional **(A)** and longitudinal **(B)** data (see Fig. 1 and main text for details). Points and error bars represent means and $\pm$ s.e., respectively.

**Figure 6** Kaplan-Meier curves showing the relationship between telomere length and survival in chick (**A**), fledgling  (**B**) and sub-adult (**C**) Seychelles warblers. Telomere length is binned into groups here for visualisation purposes only (long and short = greater than or less than median telomere length, repectively).

@@

```{r Figure 1,fig.width=6,fig.height=9}
######################
#1A
######################

@@

Fig1A <- ggplot(ddFig1A,
       aes(x = Ageclass,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'A',x=0.5,y=6.2)

@@

Fig1B <- ggplot(ddFig1B,
       aes(x = Ageclass,
           y = TROCmean)) +
  ylab('Corrected telomere loss') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  annotate('text',label = 'B',x=0.5,y=0.0018)

@@

Fig2A <- ggplot(juvseason,
       aes(x = CatchYear,
           y = TLKBmean,
           col = Season)) +
  ylab('Telomere length') + 
  xlab('Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse),
                  position = position_dodge(width = 0.2)) +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'A',x=1995,y=10)

@@

Fig2B <- ggplot(juvseason,
       aes(x = Insect,
           y = TLKBmean,
           col = Season)) +
  ylab('Telomere length') + 
  xlab('Food availability') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  stat_smooth(method = 'lm',se=T,col = 'darkgrey') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'B',x=-1.5,y=8)

@@

temp <- loss.season[complete.cases(loss.season),]

Fig2C <- ggplot(temp,
       aes(x = CatchYear,
           y = TROCmean,
           col = Season)) +
  ylab('Corrected telomere loss') + 
  xlab('Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TROCmean-TROCse,ymax=TROCmean+TROCse),
                  position = position_dodge(width = 0.2)) +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'C',x=1995,y=0.005)

####################################
Fig2D <- ggplot(temp,
       aes(x = Insect,
           y = TROCmean,
           col = Season)) +
  geom_pointrange(aes(min=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  ylab('Corrected telomere loss') + 
  xlab('Food availability') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  stat_smooth(method = 'lm', se=T, col = 'darkgrey') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'D',x=-1.5,y=0.0075)

@@

```{r Figure 3,fig.width=9,fig.height=9}

@@

Fig3A <- forestplot(table2) +
  annotate('text',label = 'A',x=0.6,y=0.6)

@@

ddFig3B <- ddply(temp,
                 .(HelperF),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

Fig3B <- ggplot(ddFig3B,
       aes(x = HelperF,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'B',x=0.5,y=8)

@@

Fig3C <- ggplot(chicks,
                aes(x = Tarsus,
                    y = TLKB)) +
  ylab('Telomere length (kb)') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  geom_point(size=3) +
  stat_smooth(method = 'lm',col = 'black') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'C',x=19,y=15)

multiplot(Fig3A,Fig3B,Fig3C,layout=matrix(c(1,1,2,3),2,2,byrow=TRUE))

@@

```{r Figure 5,fig.height=8,fig.width=4}

Fig5A <- ggplot(juvseason,
                aes(x = TLKBmean,
                    y = Lifespan,
                    col = Season)) +
  geom_pointrange(aes(ymin=Lifespan-Lifespanse,ymax=Lifespan+Lifespanse)) +
  geom_errorbarh(aes(xmin = TLKBmean-TLKBse,xmax = TLKBmean+TLKBse)) +
  theme_classic() +
  ylab('Remaining life') + 
  xlab('Telomere length (kb)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',x=2.5,y=12.5,label='A')

Fig5B <- ggplot(loss.season,
                aes(x = TROCmean,
                    y = Lifespan,
                    col = Season)) +
  geom_pointrange(aes(ymin=Lifespan-Lifespanse,ymax=Lifespan+Lifespanse)) +
  geom_errorbarh(aes(xmin = TROCmean-TROCse,xmax = TROCmean+TROCse)) +
  theme_classic() +
  ylab('Remaining life') + 
  xlab('Corrected telomere loss') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 11), 
        axis.text.y = element_text(size = 11), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none') +
  scale_color_manual(values = wes_palette("Royal1")) +
  stat_smooth(method = 'lm',col = 'darkgrey',fullrange = T) +
  annotate('text',x=-0.004,y=14,label='B')

@@

**Figure 6**

```{r Figure 6,fig.height=9,fig.width=4}

s1 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = chicks)
Fig6A <- ggsurv(s1,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='A')

###########################################
s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = subset(FlSAall,Ageclass == 'FL'))
Fig6B <- ggsurv(s2,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none',
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='B')

s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = subset(FlSAall,Ageclass == 'SA'))
Fig6C <- ggsurv(s2,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none',
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='C')

multiplot(Fig6A,Fig6B,Fig6C)
